Distributional Trees and Forests

### 2018-07-15 pkg/disttree/inst/JSM2018/slides.Rnw
slides for JSM 2018
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2219 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 2c56fe0e7f2977f91d1dc0720277c2e07b385288

\title{Flexible Probabilistic Forecasting with Distributional Trees and Forests}
\author{Lisa Schlosser, Torsten Hothorn, Reto Stauffer, Achim Zeileis}
%\author{Lisa Schlosser, Torsten Hothorn,\\ Heidi Seibold, Reto Stauffer, Achim Zeileis}
\setbeamerfont{url}{size*={11.5pt}{13pt},series=\mdseries}
\URL{https://R-Forge.R-project.org/projects/partykit/}
\headerimage{3}

\setbeamertemplate{caption}{\insertcaption} 
%% includes a replacement for \usepackage{Sweave}
\usepackage{changepage}
\usepackage{amsmath,tikz}
\usepackage{calc}
\usepackage{graphicx}
\usetikzlibrary{positioning,shapes,arrows,decorations.pathreplacing,calc,automata,mindmap,trees,tikzmark,decorations.pathreplacing}
\graphicspath{{plots/}}
\newcommand{\argmax}{\operatorname{argmax}\displaylimits}

@@

# setwd("~/svn/partykit/pkg/disttree/inst/eeecon_workshop")
library("ggplot2")
theme_set(theme_bw(base_size = 18))
library("disttree")
#library("partykitR1")
library("gamlss")
library("lattice")
library("crch")
library("latex2exp")
library("parallel")
library("RainTyrol")
library("gamlss.cens")
gen.cens(NO, type = "left")

## HCL palette
pal <- hcl(c(10, 128, 260), 100, 50)
names(pal) <- c("forest", "tree", "gamlss")

pallight <- hcl(c(10, 128, 260), 100, 50, alpha = 0.25)
names(pallight) <- c("forest", "tree", "gamlss")

@@

######## data generating process based on a predefined tree or a parameter function
# if tree: 2 splits => 3 terminal nodes
# 10 split variables (x1, ..., x10) are available (5 numeric, 2 binary, 3 categorical)
#
# input: n            ..... nr of observations
#        family       ..... distribution of the generated observations
#        split.matrix ..... indices of the variables used for the splits together with the corresponding split points
#        parm.matrix  ..... set of distribution parameters for each subgroup
#        fun          ..... parameter function
#
# output: data.frame with generated observations y (generated seperatly in each subgroups with given distributions parameters),
#                         the given split variables x1, ..., x10 for each observation and
#                         the index of the subgroup for each observation or
#                         the distribution parameter for each observation

@@

\begin{frame}[fragile]
\frametitle{Statistical Modeling}
\visible<2,3, 5-> {Assumption:}
\vspace{-0.7cm}
\begin{center}
\begin{tikzpicture}
\visible<2,3, 5-> {\node[] at (0,0.4) {X};}
\visible<4>{
\node[inner sep=0pt] (today) at (-1.8,-0.4)
    {\includegraphics[width=.365\textwidth]{today.png}};
\node[font=\scriptsize] at (-1.9,-2) {2018-03-15};
}
\visible<2->{
\draw[->] (0.4,0.4)--(0.9,0.4);
\draw[] (1,0) rectangle (2.5,0.8);
\node[font=\scriptsize] at (1.75,0.4) {nature};
\draw[->] (2.6,0.4)--(3.1,0.4);
}
\visible<2,3, 5->{
\node[] at (3.5,0.4) {Y};
}
\visible<4>{
\node[inner sep=0pt] (future) at (5.3,-0.4)
    {\includegraphics[width=.365\textwidth]{future.png}};
\node[font=\scriptsize] at (5.4,-2) {2018-03-16};
}
\end{tikzpicture}
\end{center}
\visible<3, 5->{
\vspace{-2cm}
Approximation of reality based on a set of observations.\\
}
\visible<4>{
\vspace{2cm}
X = observed precipitation of today,\\
\qquad current temperature / humidity / \dots,\\
\qquad physical model output\\
\vspace{0.15cm}
Y = daily precipitation sums\\
}

\vspace{-3.5cm}
\visible<6->{
\textbf{Application/Motivation:}
\begin{itemize}
\item Prediction.
\item Capturing dependencies and interactions.
\end{itemize}
}
\vspace{0.5cm}
\visible<7->{
\textbf{Two main approaches / ``cultures'':}
\begin{itemize}
\item Data modeling.
\item Algorithmic modeling.
\end{itemize}
}
\end{frame}

\begin{frame}[fragile]
\frametitle{Data Modeling}
\vspace{-0.1cm}
\begin{center}
\begin{tikzpicture}
\node[] at (0,0.4) {X};
\draw[->] (0.4,0.4)--(0.9,0.4);
\draw[] (1,0) rectangle (2.5,0.8);
\node[font=\scriptsize] at (1.75,0.4) {f(X; par)};
\draw[->] (2.6,0.4)--(3.1,0.4);
\node[] at (3.5,0.4) {Y};
\end{tikzpicture}
\end{center}
\vspace{-0.4cm}
\only<1-5>{
\begin{itemize}
\item Approximate data generating process up to a few unknown processes.
\item Mathematical equations, model assumptions.
\item Data: estimation of parameters.
\end{itemize}
}
\only<6->{
\vspace{1cm}
Summary:
\begin{itemize}
\item Stochastic model to approximate data generating process.\\
\only{$\Rightarrow$ fixed $f$.\\}
\item Disadvantage: Risk of incorrect assumptions.
\end{itemize}
}
\only<1>{\vspace{1cm}}

\vspace{0.1cm}

\visible<2-5>{
\begin{minipage}{0.65\textwidth}
\textbf{Example:} 
\only<2-3>{Linear model}\only<4>{Generalized additive model}\only<5-6>{Generalized additive model\\ for location, scale, and shape}
\\
\only<2-3>{
$$\visible<3>{\mathbb{E}(Y|X) = } f(X; \alpha) = \alpha_0 + \alpha_1 \cdot X$$  
\vspace{0.15cm}
}
\only<4>{
$$\mathbb{E}(Y|X) = f(X; \alpha) = \alpha + s(X)$$  
\vspace{0.15cm}
}
\only<5>{
\vspace{-1.29cm}

\begin{align*}
\visible<5>{\mathbb{E}(Y|X) = }\mu(X) = \alpha_{\mu} + s_{\mu}(X)\\  
\visible<5>{\sqrt{Var(Y|X)} = }\sigma(X) = \alpha_{\sigma} + s_{\sigma}(X)
\end{align*}
}
\end{minipage}
%\vspace{-0.6cm}
%\hfill
\begin{minipage}{0.3\textwidth}
<<motivation_GLM, echo=FALSE, results=hide>>=
nobs <- 200
## GLM
set.seed(7)
x <- c(1:nobs)/nobs
ytrue <- 1+x*1.5
y <- ytrue + rnorm(nobs,0,0.3)
@
\only<2-3>{
<<plot_motivation_GLM_single, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(y=y , x=x, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytrue, col = pal["forest"], lwd = 7, main = "")
@
}
<<motivation_GAM, echo=FALSE, results=hide>>=
## GAM
set.seed(7)
x <- c(1:nobs)/nobs
x <- 2*(x-0.5)
ytrue <- x^3 
y <- ytrue + rnorm(nobs,0,0.3)
@
\only<4>{
<<plot_motivation_GAM, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(y=y , x=x, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytrue, col = pal["forest"], lwd = 7, main = "")
@
}
<<motivation_GAMLSS, echo=FALSE, results=hide>>=
## GAMLSS
set.seed(7)
x <- c(1:nobs)/nobs
x <- 2*(x-0.5)
ytrue <- x^3
var <- exp(-(2*x)^2)/2
y <- ytrue + rnorm(nobs, 0, 0.1 + var)
@
\only<5>{
<<plot_motivation_GAMLSS, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x, y, xaxt = "n", yaxt = "n", ann = FALSE, type = "n")
polygon(c(x, rev(x)), c(ytrue + 0.1 + var, rev(ytrue - 0.1 - var)),
  col = pallight["forest"], border = "transparent")
lines(x, ytrue, col = pal["forest"], lwd=7)
points(x, y, col = "slategray", pch = 19)
box(lwd = 5)
@
}
\end{minipage}
}
\end{frame}

<<raintree, echo=FALSE, results=hide>>=
load("weather.rda")
#d <- data.frame(rain24 = numeric(0), rain = logical(0), wind = numeric(0), 
#                humidity = numeric(0), airpressure = numeric(0), temperature = numeric(0))

sp_rain <- partysplit(2L, index = 1:2)
sp_humidity <- partysplit(4L, breaks = 70)

pn <- partynode(1L, split = sp_rain, kids = list(
  partynode(2L, split = sp_humidity, kids = list(
      partynode(3L, info = "4 mm/24h"),
      partynode(4L, info = "9 mm/24h"))),
  partynode(5L, info = "15 mm/24h")))

p <- party(pn, weather)
@
<<motivation_regtree, echo=FALSE, results=hide>>=
## Reg. Tree
set.seed(7)
kappa <- 12
nobs <- 120
x <- c(1:nobs)/nobs
ytrue <- ytree <- ytree2 <- ytree3 <- ytree4 <- 
  ytree_up <- ytree_up2 <- ytree_up3 <- ytree_up4 <- 
  yforest <- numeric(length = length(x))
for(i in 1:nobs) ytrue[i] <- if(x[i]<0.55) 4 else 4+5*plogis(kappa*(3*(x[i]-0.36)-1))
y <- ytrue + rnorm(nobs,0,2)
y_up <- rep.int(15, length(x)) + rnorm(nobs,0,1)
for(i in 1:nobs) ytree[i] <- if(x[i]<0.7) 4 else 9
for(i in 1:nobs) ytree2[i] <- if(x[i]<0.63) 4.3 else 9.4
for(i in 1:nobs) ytree3[i] <- if(x[i]<0.75) 3.7 else 9.2
for(i in 1:nobs) ytree4[i] <- if(x[i]<0.55) 4.4 else 8.7
ytree_up <- rep.int(15, length(x))
ytree_up2 <- rep.int(14.5, length(x))
ytree_up3 <- rep.int(15.3, length(x))
ytree_up4 <- rep.int(15.5, length(x))
for(i in 1:nobs) yforest[i] <- if(x[i]<0.55) 4 else {if(x[i]<0.63) 5 else {if(x[i]<0.7) 6.5 else {if(x[i]<0.75) 8 else 9}}}
@
<<motivation_randforest, echo=FALSE, results=hide>>=
## Random Forest
#%for(i in 1:nobs) yforest[i] <- if(x[i]<0.27) 0.5 else { if(x[i]<0.39) 0.5 + 1.5*(plogis((x[i]-0.33)/6*700)) else 1+(1-plogis(kappa*(2*(x[i]-0.2)-1)))}
@

\begin{frame}[fragile]
\frametitle{Algorithmic Modeling}
\vspace{-0.1cm}
\begin{center}
\begin{tikzpicture}
\node[] at (0,0.4) {X};
\draw[->] (0.4,0.4)--(0.9,0.4);
\draw[] (1,0) rectangle (2.5,0.8);
\node[font=\scriptsize] at (1.75,0.4) {nature};
\draw[->] (2.6,0.4)--(3.1,0.4);
\node[] at (3.5,0.4) {Y};
\draw[<->] (0.1,0.1) .. controls (0.2, -0.4) and (0.4, -0.5) .. (0.8,-0.5);
\draw[<->] (2.7,-0.5) .. controls (3.1, -0.5) and (3.3, -0.4) .. (3.4,0.1);
\node[font=\scriptsize] at (1.75,-0.5) {algorithm};
\end{tikzpicture}
\end{center}
\vspace{-0.4cm}
\visible<2->{
\begin{itemize}
\item Flexible approximation with no/few assumptions. %\\
%$\Rightarrow$ no fixed $f$.
\item Data-driven model fitting.
\end{itemize}
}
\only<2>{\vspace{1cm}}
\visible<3->{
\begin{minipage}{0.6\textwidth}
%\vspace{-0.5cm}
\textbf{Example:} \only<3>{Tree model}\only<4-10>{Subsampling}\only<3>{\\}\only<11->{Forest model}\\
%\vspace{-0.8cm}
\begin{center}
\only<3>{
\vspace{-0.96cm}
\resizebox{0.6\textwidth}{!}{
<<plot_raintree, fig=TRUE, echo=FALSE, width=4, height=2.7>>=
par(mar=c(0,0,0,0), oma=c(0,0,0.4,0))
paltrees <- rgb(c(0.97, 0.64, 1), c(0.70, 0.83, 1), c(0.30, 0.99, 1))
plot(p,
     tp_args = list(FUN = identity, width = 9, fill = paltrees[c(1, 3)]), 
     ip_args = list(fill = paltrees[c(2, 3)]),
     drop_terminal = TRUE, tnex = 1)
@
}

\vspace{-0.7cm}
\begin{flushleft}
\noindent {\small X = rain now, humidity\\
Y = rain 24h}
\end{flushleft}}
\only<4->{
\vspace{-0.5cm}
\resizebox{0.55\textwidth}{!}{
\begin{tikzpicture}
%\draw[ellipse, draw=black, align=center, scale = 0.7, minimum width=170pt, minimum height = 35pt] (set) at (1, 3) {};
\draw[gray,rounded corners=10pt] (-1,4.3) -- (4,4.3) -- (4,2.7) -- (-2,2.7) -- (-2,4.3) -- (0,4.3);
%\draw[gray] (1,3.5) ellipse (3 and 0.8);
\begin{scope}
\clip (-1,4.3) -- (4,4.3) -- (4,2.7) -- (-2,2.7) -- (-2,4.3) -- (0,4.3);
%\clip (1,3.5) ellipse (3 and 0.8);
\pgfmathsetseed{7}
\foreach \p in {1,...,200} {\fill[gray] (1+3*rand,3.5+0.8*rand) circle (0.04);}
\end{scope}
\visible<5->{
\node[ellipse, draw=black!50!green, align=center, minimum width=80pt, minimum height = 40pt, line width = 3pt] (subset1) at (0.1, 3.5) {};
}
\visible<7->{
\node[ellipse, draw = blue, align=center, minimum width=100pt, minimum height = 30pt, line width = 3pt] (subset2) at (1.2, 3.6) {};
}
\visible<9->{
\node[ellipse, draw = orange, align=center, minimum width=120pt, minimum height = 25pt, line width = 3pt] (subset3) at (1.6, 3.3) {};
}
\visible<6->{
  \node[ellipse, fill=black!50!green, align=center] (n00) at (-1, 2) {};
  \node[ellipse, fill=black!50!green, align=center] (n01) at (-1.25, 1.25) {};
  \draw[-, line width=1pt] (n00) -- (n01);
  \node[rectangle, fill=black!50!green, align=center] (n02) at (-1.5, 0.5) {};
  \draw[-, line width=1pt] (n01) -- (n02);
  \node[rectangle, fill=black!50!green, align=center] (n03) at (-1, 0.5) {};
  \draw[-, line width=1pt] (n01) -- (n03);
  \node[rectangle, fill=black!50!green, align=center] (n04) at (-0.5, 0.5) {};
  \draw[-, line width=1pt] (n00) -- (n04);
}
\visible<8->{
  \node[ellipse, fill=blue, align=center] (n10) at (1, 2) {};
  \node[ellipse, fill=blue, align=center] (n11) at (0.75, 1.25) {};
  \draw[-, line width=1pt] (n10) -- (n11);
  \node[rectangle, fill=blue, align=center] (n12) at (0.5, 0.5) {};
  \draw[-, line width=1pt] (n11) -- (n12);
  \node[rectangle, fill=blue, align=center] (n13) at (1, 0.5) {};
  \draw[-, line width=1pt] (n11) -- (n13);
  \node[rectangle, fill=blue, align=center] (n14) at (1.5, 0.5) {};
  \draw[-, line width=1pt] (n10) -- (n14);
}
\visible<10->{
  \node[ellipse, fill=orange, align=center] (n20) at (3, 2) {};
  \node[ellipse, fill=orange, align=center] (n21) at (2.75, 1.25) {};
  \draw[-, line width=1pt] (n20) -- (n21);
  \node[rectangle, fill=orange, align=center] (n22) at (2.5, 0.5) {};
  \draw[-, line width=1pt] (n21) -- (n22);
  \node[rectangle, fill=orange, align=center] (n23) at (3, 0.5) {};
  \draw[-, line width=1pt] (n21) -- (n23);
  \node[rectangle, fill=orange, align=center] (n24) at (3.5, 0.5) {};
  \draw[-, line width=1pt] (n20) -- (n24);
}
\visible<11->{
\draw[-, line width=1pt] (-1, -0.2) -- (0.2,-0.7);
\draw[-, line width=1pt] (1, -0.2) -- (1,-0.7);
\draw[-, line width=1pt] (3, -0.2) -- (1.8,-0.7);
\node[rectangle, align=center] (combine) at (1, -1) {\large average};
}
\end{tikzpicture}
}
}
\end{center}
\end{minipage}
\hspace{-1cm}
\begin{minipage}{0.4\textwidth}
\begin{center}
\vspace{0.1cm}

\only<3-5>{
%\vspace{0.01cm}
<<plot_motivation_regtree, fig=TRUE, echo=FALSE, width=7>>=
par(mar=c(3,3,2,5.5))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19,
     ylim = c(min(c(y,y_up)), max(c(y,y_up))))
points(x=x, y = y_up, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytree, col = pal["forest"], lwd=7)
lines(x = x, y = ytree_up, col = pal["forest"], lwd=7)
mtext("rain", side = 4, las = TRUE, at = 16, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("now", side = 4, las = TRUE, at = 14.7, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("no", side = 4, las = TRUE, at = 10.3, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("rain", side = 4, las = TRUE, at = 9, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("now", side = 4, las = TRUE, at = 7.7, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("rain 24h", side = 2, las = FALSE, line = 1, cex = 3)
mtext("humidity", side = 1, line = 1.5, cex = 3)
\only<6-7>{
%\vspace{0.01cm}
\hspace{-0.2cm}
<<plot_motivation_randforest1, fig=TRUE, echo=FALSE, width=7>>=
par(mar=c(3,3,2,5.5))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19,
     ylim = c(min(c(y,y_up)), max(c(y,y_up))))
points(x=x, y = y_up, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytree, col = pal["forest"], lwd=7)
lines(x = x, y = ytree_up, col = pal["forest"], lwd=7)
lines(x = x, y = ytree2, col = pal["tree"], lwd=7)
lines(x = x, y = ytree_up2, col = pal["tree"], lwd=7)
mtext("rain", side = 4, las = TRUE, at = 16, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("now", side = 4, las = TRUE, at = 14.7, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("no", side = 4, las = TRUE, at = 10.3, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("rain", side = 4, las = TRUE, at = 9, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("now", side = 4, las = TRUE, at = 7.7, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("rain 24h", side = 2, las = FALSE, line = 1, cex = 3)
mtext("humidity", side = 1, line = 1.5, cex = 3)
\only<8-9>{
%\vspace{0.01cm}
\hspace{-0.3cm}
<<plot_motivation_randforest2, fig=TRUE, echo=FALSE, width=7>>=
par(mar=c(3,3,2,5.5))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19,
     ylim = c(min(c(y,y_up)), max(c(y,y_up))))
points(x=x, y = y_up, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytree, col = pal["forest"], lwd=7)
lines(x = x, y = ytree_up, col = pal["forest"], lwd=7)
lines(x = x, y = ytree2, col = pal["tree"], lwd=7)
lines(x = x, y = ytree3, col = "blue", lwd=7)
lines(x = x, y = ytree_up2, col = pal["tree"], lwd=7)
lines(x = x, y = ytree_up3, col = "blue", lwd=7)
mtext("rain", side = 4, las = TRUE, at = 16, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("now", side = 4, las = TRUE, at = 14.7, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("no", side = 4, las = TRUE, at = 10.3, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("rain", side = 4, las = TRUE, at = 9, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("now", side = 4, las = TRUE, at = 7.7, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("rain 24h", side = 2, las = FALSE, line = 1, cex = 3)
mtext("humidity", side = 1, line = 1.5, cex = 3)
\only<10>{
%\vspace{0.01cm}
\hspace{-0.3cm}
<<plot_motivation_randforest3, fig=TRUE, echo=FALSE, width=7>>=
par(mar=c(3,3,2,5.5))

>>>>

\title{Distributional Trees and Forests}
\author{Lisa Schlosser, Torsten Hothorn,\\ Heidi Seibold, Reto Stauffer, Achim Zeileis}
\setbeamerfont{url}{size*={11.5pt}{13pt},series=\mdseries}
%\renewcommand{\mysize}{\fontsize{7.7pt}{9pt}\selectfont}
\URL{https://R-Forge.R-project.org/projects/partykit/}
\headerimage{3}

\setbeamertemplate{caption}{\insertcaption} 
%% includes a replacement for \usepackage{Sweave}
\usepackage{amsmath,tikz}
\usetikzlibrary{positioning,shapes,arrows,decorations.pathreplacing,calc,automata,mindmap}
\graphicspath{{plots/}}

\newcommand{\argmax}{\operatorname{argmax}\displaylimits}

@@

# setwd("~/svn/partykit/pkg/disttree/inst/JSM2018")
library("ggplot2")
theme_set(theme_bw(base_size = 18))
library("disttree")
library("Formula")
#library("partykitR1")
library("gamlss")
library("lattice")
library("crch")
library("latex2exp")
library("parallel")
library("gamlss.cens")
gen.cens(NO, type = "left")

## HCL palette
pal <- hcl(c(10, 128, 260, 290, 50), 100, 50)
names(pal) <- c("forest", "tree", "gamlss", "gamboostLSS", "EMOS")

pallight <- hcl(c(10, 128, 260, 290, 70), 100, 50, alpha = 0.25)
names(pallight) <- c("forest", "tree", "gamlss", "gamboostLSS", "EMOS")

@@

######## data generating process based on a predefined tree or a parameter function
# if tree: 2 splits => 3 terminal nodes
# 10 split variables (x1, ..., x10) are availabe (5 numeric, 2 binary, 3 categorical)
#
# input: n            ..... nr of observations
#        family       ..... distribution of the generated observations
#        split.matrix ..... indices of the variables used for the splits together with the corresponding split points
#        parm.matrix  ..... set of distribution parameters for each subgroup
#        fun          ..... parameter function
#
# output: data.frame with generated observations y (generated seperatly in each subgroups with given distributions parameters),
#                         the given split variables x1, ..., x10 for each observation and
#                         the index of the subgroup for each observation or
#                         the distribution parameter for each observation

@@

\subsection{Motivation}

\begin{frame}[fragile]
\frametitle{Motivation}
\vspace{-0.41cm}
\begin{figure}[!htb]
\minipage{0.29\textwidth}
\begin{center}
<<motivation_GLM, echo=FALSE, results=hide>>=
nobs <- 200
## GLM
set.seed(7)
x <- c(1:nobs)/nobs
ytrue <- 1+x*1.5
y <- ytrue + rnorm(nobs,0,0.3)
@
\visible<2->{
<<plot_motivation_GLM, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(y=y , x=x, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytrue, col = pal["forest"], lwd = 7, main = "")
@
}
\end{center}
\endminipage
\visible<3->{{\LARGE$\rightarrow$}}
\minipage{0.29\textwidth}
\begin{center}
<<motivation_GAM, echo=FALSE, results=hide>>=
## GAM
set.seed(7)
x <- c(1:nobs)/nobs
x <- 2*(x-0.5)
ytrue <- x^3 
y <- ytrue + rnorm(nobs,0,0.3)
@
\visible<3->{
<<plot_motivation_GAM, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(y=y , x=x, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytrue, col = pal["forest"], lwd = 7, main = "")
@
}
\end{center}
\endminipage
\visible<4->{{\LARGE$\rightarrow$}}
\minipage{0.29\textwidth}
\begin{center}
<<motivation_GAMLSS, echo=FALSE, results=hide>>=
## GAMLSS
set.seed(7)
x <- c(1:nobs)/nobs
x <- 2*(x-0.5)
ytrue <- x^3
var <- exp(-(2*x)^2)/2
y <- ytrue + rnorm(nobs, 0, 0.1 + var)
@
\visible<4->{
<<plot_motivation_GAMLSS, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x, y, xaxt = "n", yaxt = "n", ann = FALSE, type = "n")
polygon(c(x, rev(x)), c(ytrue + 0.1 + var, rev(ytrue - 0.1 - var)),
  col = pallight["forest"], border = "transparent")
lines(x, ytrue, col = pal["forest"], lwd=7)
points(x, y, col = "slategray", pch = 19)
box(lwd = 5)
@

}
\end{center}
\endminipage

\vspace{0.5cm}
\minipage{0.25\textwidth}
\begin{center}
\visible<2->{
LM, GLM\\
\vspace{0.5cm}
\code{lm}\\
\code{glm}\\
\vspace{1.5cm}}
\end{center}
\endminipage
\hspace{1.1cm}
\minipage{0.25\textwidth}
\begin{center}
\visible<3->{
GAM\\
\vspace{0.5cm}
\code{mgcv}\\
\code{VGAM}\\
$\ldots$\\
\vspace{1cm}}
\end{center}
\endminipage
\hspace{1.1cm}
\minipage{0.25\textwidth}
\begin{center}
\visible<4->{
GAMLSS\\
\vspace{0.5cm}
\code{gamlss}\\
\code{mgcv}\\
\code{VGAM}\\
\code{gamboostLSS}\\
$\ldots$}
\end{center}
\endminipage
\end{figure}
\end{frame}

\begin{frame}[fragile]
\frametitle{Motivation}
\vspace{-0.2cm}
\begin{figure}[!htb]
\minipage{0.29\textwidth}
\begin{center}
<<motivation_regtree, echo=FALSE, results=hide>>=
## Reg. Tree
set.seed(7)
kappa <- 12
x <- c(1:nobs)/nobs
ytrue <- ytree <- yforest <- numeric(length = length(x))
for(i in 1:nobs) ytrue[i] <- if(x[i]<1/3) 0.5 else 1+(1-plogis(kappa*(2*(x[i]-0.2)-1)))
y <- ytrue + rnorm(nobs,0,0.3)
for(i in 1:nobs) ytree[i] <- if(x[i]<1/3) 0.5 else {if(x[i]<2/3) 2 else 1}
@
\visible<1->{
<<plot_motivation_regtree, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
#lines(x = x, y = ytrue, col = "grey", lwd=5, main = "")
lines(x = x, y = ytree, col = pal["forest"], lwd=7)
@
}
\end{center}

\endminipage
\visible<2->{{\LARGE$\rightarrow$}}
\minipage{0.29\textwidth}
 \begin{center}
<<motivation_randforest, echo=FALSE, results=hide>>=
## Random Forest
for(i in 1:nobs) yforest[i] <- if(x[i]<0.27) 0.5 else { if(x[i]<0.39) 0.5 + 1.5*(plogis((x[i]-0.33)/6*700)) else 1+(1-plogis(kappa*(2*(x[i]-0.2)-1)))}
@
\visible<2->{
<<plot_motivation_randforest, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
#lines(x = x, y = ytrue, col = "grey", lwd=5, main = "")
lines(x = x, y = yforest, col = pal["forest"], lwd=7, main = "")
@
}
\end{center}

\endminipage
\visible<3->{{\LARGE$\rightarrow$}}
\minipage{0.29\textwidth}
  \visible<3->{
  \begin{center}
  \begin{tikzpicture}
  \draw[line width=0.2mm, black] (0,0) rectangle (2.6,2.24);
  \node[rectangle, align=center] (sq) at (1.3, 1.12) {\Huge{?}};
  \end{tikzpicture}
  \end{center}}
\endminipage

\minipage{0.29\textwidth}
\begin{center}
\vspace{0.0cm}
\visible<1->{
Regression Tree\\
\vspace{0.5cm}
\resizebox{0.3\textwidth}{!}{
\begin{tikzpicture}
  \node[ellipse, fill=HighlightBlue!70, align=center] (n0) at (1, 2) {};
  \node[rectangle, fill=HighlightOrange!70, align=center] (n1) at (0.5, 1) {};
  \draw[-, line width=1pt] (n0) -- (n1);
  \node[ellipse, fill=HighlightBlue!70, align=center] (n2) at (1.5, 1) {};
  \draw[-, line width=1pt] (n0) -- (n2);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n3) at (1, 0) {};
  \draw[-, line width=1pt] (n2) -- (n3);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n4) at (2, 0) {};
  \draw[-, line width=1pt] (n2) -- (n4);
\end{tikzpicture}}
\vspace{0.3cm}\\
\code{rpart}\\
\code{party(kit)}\\
\vspace{1cm}}
\end{center}
\endminipage
\hspace{0.65cm}
\minipage{0.29\textwidth}
\begin{center}
\vspace{0.0cm}
\visible<2->{
Random Forest\\
\vspace{0.5cm}
\resizebox{0.9\textwidth}{!}{
\begin{tikzpicture}
  \node[ellipse, fill=HighlightBlue!70, align=center] (n00) at (1, 2) {};
  \node[rectangle, fill=HighlightOrange!70, align=center] (n01) at (0.5, 1) {};
  \draw[-, line width=1pt] (n00) -- (n01);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n02) at (1.5, 1) {};
  \draw[-, line width=1pt] (n00) -- (n02);

  \node[ellipse, fill=HighlightBlue!70, align=center] (n10) at (3, 2) {};
  \node[ellipse, fill=HighlightBlue!70, align=center] (n11) at (2.5, 1) {};
  \draw[-, line width=1pt] (n10) -- (n11);
  \node[ellipse, fill=HighlightBlue!70, align=center] (n12) at (3.5, 1) {};
  \draw[-, line width=1pt] (n10) -- (n12);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n13) at (2, 0) {};
  \draw[-, line width=1pt] (n11) -- (n13);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n14) at (2.8, 0) {};
  \draw[-, line width=1pt] (n11) -- (n14);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n15) at (3.2, 0) {};
  \draw[-, line width=1pt] (n12) -- (n15);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n16) at (4, 0) {};
  \draw[-, line width=1pt] (n12) -- (n16);

  \node[ellipse, fill=HighlightBlue!70, align=center] (n20) at (5, 2) {};
  \node[rectangle, fill=HighlightOrange!70, align=center] (n21) at (4.5, 1) {};
  \draw[-, line width=1pt] (n20) -- (n21);
  \node[ellipse, fill=HighlightBlue!70, align=center] (n22) at (5.5, 1) {};
  \draw[-, line width=1pt] (n20) -- (n22);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n23) at (5, 0) {};
  \draw[-, line width=1pt] (n22) -- (n23);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n24) at (6, 0) {};
  \draw[-, line width=1pt] (n22) -- (n24);
\end{tikzpicture}
}
\vspace{0.5cm}\\
\code{randomForest}\\
\code{ranger}\\
\code{party(kit)}\\
$\ldots$}
\end{center}
\endminipage
\hspace{0.65cm}
\minipage{0.28\textwidth}
\begin{center}
\vspace{0.68cm}
\visible<3->{
Distributional trees and forests\\
\vspace{1.7cm}
\code{disttree}\\
based on \code{partykit}\\
\vspace{1cm}}
\end{center}
\endminipage
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Goals}

\textbf{Distributional:} \\
\begin{itemize}
\item Specify the complete probability distribution (including location, scale and shape).
\end{itemize}

{\bf Tree:}
\begin{itemize}
  \item Automatic variable selection and detection of steps and abrupt changes. %(data driven)
  \item Capture non-linear and non-additive effects and interactions.
\end{itemize}

{\bf Forest:}
\begin{itemize}
  \item Smoother effects.
  \item Stabilization and regularization of the model.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Distributional trees}
\vspace*{-0.12cm}
\begin{center} 
DGP:  $\; Y\ |\ X = x \; \sim  \; \mathcal{N}(\mu(x), \sigma^2(x))$
\vspace*{-0.21cm}
<<dgp_tree, echo=FALSE, results=hide>>=
data <- data.frame(x = numeric(0), x = numeric(0), x = numeric(0))
names(data) <- c("x","x","x")
fig <- party(
  partynode(1L,
            split = partysplit(2L, breaks = 0.4),
            kids = list(
              partynode(2L, info = c(
                "n = 200",
                "   True parameters:   ",
                expression(mu == '4'),
                expression(sigma == '1')
              )),
              partynode(3L,
                        split = partysplit(3L, breaks = 0.8),
                        kids = list(
                          partynode(4L, info = c(
                            "n = 200",
                            "   True parameters:   ",
                            expression(mu == '12'),
                            expression(sigma == '3')
                          )),
                          partynode(5L, info = c(
                            "n = 100",
                            "   True parameters:   ",
                            expression(mu == '4'),
                            expression(sigma == '3')
                          )))))),
  data
)
